---
title: "Fatal Police Shootings"
author: "Devin Luu"
output:
  html_notebook:
    code_folding: hide
    toc: true
    toc_depth: 4
    toc_float: true
runtime: shiny
---

```{r setup, include=FALSE}
library(tidyverse)
library(data.world)
library(shiny)
library(gridExtra)
library(MASS)
knitr::opts_chunk$set(echo = TRUE)
```
  
## **R Session Info**  

```{r}
sessionInfo()
```

## **Github Link** 
TBD

## **Introduction** 
Some Introduction.

## **Getting data from data.world** 
Short description of the data.
```{r}
project <- "https://data.world/devtl/fatalpoliceshootings/"
#data.world::set_config(cfg_env("DW_API"))
df <- data.world::query(data.world::qry_sql("SELECT * FROM fatalPoliceShootingsCleaned"), dataset = project)

summary(df)
```

## Possible classifications

This dataset has many qualitative variables that could do well for classification methods. I am interested in predicting a victim's threat level based on age, whether or not they are armed, age, gender, race, whether or not they are fleeing, and signs of mental illness.

```{r}
predictors <- df%>%dplyr::select("armed","age","gender","race","signs_of_mental_illness","flee")

# signs of mental illness: 1=T, 0=F
x <- factor(rep(predictors$signs_of_mental_illness))
levels(x) <- 0:1
predictors$signs_of_mental_illness = x

# fleeing: 3=car, 2=foot, 1=other, 0=not fleeing
y <- factor(rep(predictors$flee))
levels(y) = c(3,2,0,1)
predictors$flee <- y

# gender: 1=M, 0=F
w <- factor(rep(predictors$gender))
levels(w) <- 0:1
predictors$gender = w

# armed=1 or unarmed=0
z <- unlist(
lapply(
predictors$armed, function(s){ifelse(s == "unarmed", as.integer(0), as.integer(1))}
)
)
predictors <- predictors %>% mutate(armedOrUnarmed = z)
renderTable(head(predictors), hover = TRUE, width = "auto")
```

Value Mappings:
* signs of mental illness: 1=T, 0=F
* fleeing: 3=car, 2=foot, 1=other, 0=not fleeing
* gender: 1=M, 0=F
* armed=1 or unarmed=0


## Demographics

### Age
```{r}
renderPlot({
  ggplot(na.omit(df["age"]), mapping = aes(x = age)) + geom_bar(stat = "count", fill = "steelblue") + theme_minimal()
})

summary(na.omit(df$age))
```


### Gender
```{r}
renderPlot({
  pl2.1 <- ggplot(na.omit(df["gender"]), mapping = aes(x = gender, fill=gender)) + geom_bar(stat = "count", fill = "tomato2") + theme_minimal()
  pl2.2 <- ggplot(na.omit(df[c("age","gender")]), mapping = aes(x = age, fill = gender)) + geom_bar(stat = "count") + theme_minimal() + facet_wrap(~gender) + theme(legend.position="none")
  grid.arrange(pl2.1,pl2.2, ncol = 2)
})

count(df, gender)
```

### Race
```{r}
renderPlot({
  ggplot(na.omit(df["race"]), mapping = aes(x = race)) + geom_bar(stat = "count") + theme_minimal()
  ggplot(na.omit(df[c("age","race")]), mapping = aes(x = age, fill = race)) + geom_bar(stat = "count") + theme_minimal() + facet_wrap(~race) + theme(legend.position="none")

})

count(df, race)
```

## Linear Discriminant Analysis

Using 10-fold cross-validation.
```{r}
### TRAIN/TEST 
df1 <- cbind(threat = df$threat_level,predictors)
train <- sample(1:nrow(df1), 0.9*nrow(df1))
testDf <- na.omit(df1[-train,])


### LDA
lda.fit <- lda(threat ~ age + gender + flee + signs_of_mental_illness + armedOrUnarmed, data = df1, subset = train)

renderPrint(lda.fit)
```


```{r}
ldaPred <- data.frame(predict(lda.fit,testDf))

renderPlot({
  pl1 <- ggplot(ldaPred) + geom_histogram(mapping = aes(x=x.LD1), color = "steelblue") + facet_wrap(~ class) + theme_minimal()
  pl2 <- ggplot(ldaPred) + geom_histogram(mapping = aes(x=x.LD2), color = "tomato2") + facet_wrap(~ class) + theme_minimal()
  grid.arrange(pl1,pl2, nrow =2)
}, width = 'auto')

renderPlot({
  pl3 <- ggplot(ldaPred) + geom_boxplot(mapping = aes(x=class, y=x.LD1)) + theme_minimal()
  pl4 <- ggplot(ldaPred) + geom_boxplot(mapping = aes(x=class, y=x.LD2)) + theme_minimal()
  grid.arrange(pl3,pl4, nrow =2)
}, width = 'auto')
```

```{r}
renderPlot({
  ggplot(ldaPred, mapping = aes(x.LD1, x.LD2, color = class)) + geom_point(alpha = 0.5) + theme_minimal()
})

testGroup = testDf$threat
ldaClass=ldaPred$class
table(ldaClass, testGroup)
mean(ldaClass == testGroup)
```


